---
title: "COVID_NY"
author: "Reese Hirota"
date: "2/25/2021"
output: pdf_document
---
1. State - 
ny

- less than 50 mil
- publically available intervention data
- converging data



2. Plot Data of new cases and deaths - march 15th 2020 to nov 1st 2020
```{r}
library(gridExtra)
library(covidcast)
library(ggplot2)
# the covid cast package takes a start_day, end_day, cases or deaths, and geographic value
ny_cases <- covidcast_signal(data_source = "jhu-csse",
                               signal ="confirmed_incidence_num",
                               start_day = "2020-03-15", end_day = "2020-11-01",
                               geo_type = "state", geo_values = "ny")
## GGplot syntax may take some time to get used to 
ggplot(ny_cases,aes(x=time_value,y=value)) + geom_point() + theme_bw() + xlab("Date") + ylab("New Cases")
ny_deaths <- covidcast_signal(data_source = "jhu-csse",
                               signal ="deaths_incidence_num",
                               start_day = "2020-03-15", end_day = "2020-11-01",
                               geo_type = "state", geo_values = "ny")
ggplot(ny_deaths,aes(x=time_value,y=value)) + geom_point() + theme_bw() + xlab("Date") + ylab("New Deaths")

```
3. Model parameters
from website : https://github.com/midas-network/COVID-19
all except beta


4. design model on how we expect covid 19 to flow - it should have new cases and deaths
no H - ambiguity on whether its counted as infectious people or not 
```{r}
s_t <- c(19450000) #19 million
e_t <- c(0)
i_t <- c(729) #check online
d_t <- c(0)
r_t <- c(0)
gamma <- 1/12 #symptom onset to recovery (I -> R)
beta <- 2 #FIGURE IT OUT 0.2
sigma <- 1/9.55 #average of 10.69+8.41 - values of incubation in the US (E->I)
case_fatality <- 0.055 #average of symptom onset to death (I -> D)
r_0 <- beta/gamma
new_cases <- c(0)
new_deaths <- c(0)
N <- s_t[1] + i_t[1] 
for (nsim in 1:100){
  for (t in 2:231){
    p_s_i <- beta*i_t[t-1]/N
    #if (t >2 & t<7){
    #  p_s_i <- 0.85*beta*i_t[t-1]/N
    # }
    # if (t>7 & t<19){
    # p_s_i <- 0.35*beta*i_t[t-1]/N
    # }
    # if (t>19 & t<25){
    # p_s_i <- 0.02*beta*i_t[t-1]/N
    #  }
    #if (t>25){
    #  p_s_i <- 0.028*beta*i_t[t-1]/N
    # }
    if (t>2 & t<7) {
      p_s_i <- 1.98*beta*i_t[t-1]/N
    }
    if(t>7 & t<16){
      p_s_i <- 0.7*beta*i_t[t-1]/N
    }
    if(t>16){
      p_s_i <-  0.05*beta*i_t[t-1]/N
    } 
    if (t > 30 & t <40) {
      p_s_i <-  0.035*beta*i_t[t-1]/N
    }
    if (t > 50 & t <100) {
      p_s_i <-  0.03*beta*i_t[t-1]/N
    }
    if (t >150){
      p_s_i <-  .11*beta*i_t[t-1]/N
    }
    n_t_s_e <- rbinom(1,s_t[t-1],p_s_i)
    n_t_e_i <- rbinom(1,e_t[t-1],sigma)
    n_t_i_star <- rbinom(1,i_t[t-1],gamma+case_fatality)
    n_t_i_d_and_n_t_i_r <- rmultinom(1,n_t_i_star,c(case_fatality/(gamma+case_fatality),gamma/(gamma+case_fatality)))
    n_t_i_d <- n_t_i_d_and_n_t_i_r[1]
    n_t_i_r <- n_t_i_d_and_n_t_i_r[2]
    s_t[t] <- s_t[t-1] - n_t_s_e 
    e_t[t] <- e_t[t-1] - n_t_e_i + n_t_s_e
    i_t[t] <- i_t[t-1] - n_t_i_r + n_t_e_i - n_t_i_d
    d_t[t] <- d_t[t-1] + n_t_i_d 
    r_t[t] <- r_t[t-1] + n_t_i_r
    new_cases[t] <-n_t_e_i # ask if there is a way to prevent double count since someone + in e compartment would be counted as new E and when moves to I would be counted AGAIN as new I
    new_deaths[t] <- n_t_i_d
  }  
}
plot(new_cases, type ='l', col='red',ylim=c(0,40000))
lines(new_deaths, col='black')
lines(ny_cases$value, col='blue')
lines(ny_deaths$value, col = 'green')
#abline(v=7) # add lines to timesteps where interventions happened
```

5.
```{r}
simulate_expected_curve <- function(beta){
  
matrix_of_simulations <- matrix(NA,nrow=100,ncol=231)
matrix_of_deaths <- matrix(NA,nrow=100,ncol=231)

for (nsim in 1:100){
s_t <- c(19450000) #check online
e_t <- c(0)
i_t <- c(729) #check online
d_t <- c(0)
r_t <- c(0)

gamma <- 1/12 #symptom onset to recovery (I -> R)
beta <- 2 #FIGURE IT OUT 0.2
sigma <- 1/9.55 #average of 10.69+8.41 - values of incubation in the US (E->I)
case_fatality <- 0.06 #average of symptom onset to death (I -> D)
r_0 <- beta/gamma
new_cases <- c(0)
new_deaths <- c(0)

N <- s_t[1] + i_t[1]
  
  for (t in 2:231){
    
    p_s_i <-  beta*i_t[t-1]/N
    
    n_t_s_e <- rbinom(1,s_t[t-1],p_s_i)
    n_t_e_i <- rbinom(1,e_t[t-1],sigma)
    n_t_i_star <- rbinom(1,i_t[t-1],min(1, gamma+case_fatality))
    n_t_i_d_and_n_t_i_r <- rmultinom(1,n_t_i_star,c(case_fatality/(gamma+case_fatality),gamma/(gamma+case_fatality)))
    n_t_i_d <- n_t_i_d_and_n_t_i_r[1]
    n_t_i_r <- n_t_i_d_and_n_t_i_r[2]
  
   s_t[t] <- s_t[t-1] - n_t_s_e 
   e_t[t] <- e_t[t-1] - n_t_e_i + n_t_s_e
   i_t[t] <- i_t[t-1] - n_t_i_r + n_t_e_i - n_t_i_d
   d_t[t] <- d_t[t-1] + n_t_i_d 
   r_t[t] <- r_t[t-1] + n_t_i_r
   new_cases[t] <- n_t_e_i # ask if there is a way to prevent double count since someone + in e compartment would be counted as new E and when moves to I would be counted AGAIN as new I
   
   new_deaths[t] <- n_t_i_d
    }
  
  matrix_of_simulations[nsim,] <- new_cases
  matrix_of_deaths[nsim,] <- new_deaths
  }
return (colMeans(matrix_of_simulations))
return(colMeans(matrix_of_deaths))
}
truth <- ny_cases$new_cases # not sure 
guess <- simulate_expected_curve(2) #change value of beta
plot(truth,type='l',ylim=c(0,1000),xlim=c(0,1000),col='red')
lines(guess)
print (mean((truth-guess)^2))

truth2 <- ny_deaths$new_deaths
guess2 <- simulate_expected_curve(2) #change value of beta
plot(truth2,type='l',ylim=c(0,1000),xlim=c(0,1000),col='black')
lines(guess2)
print (mean((truth2-guess2)^2))

```
```{r}
head(ny_cases)
```

6. forecast model:
a) another month - change to 261
b) end of pandemic - change t to something real big like 900
```{r}
#S -> E -> I -> R 
#           `> D


s_t <- c(0) #check online
e_t <- c(0)
i_t <- c(1000) #check online
d_t <- c(0)
r_t <- c(0)

gamma <- 1/12 #check online
beta <- 2  #check online 
sigma <- 1/9.55 #check online
r_0 <- beta/gamma
new_cases <- c(0)
new_deaths <- c(0)

N <- s_t[1] + i_t[1] 

for (nsim in 1:100){
for (t in 2:231){ #change this to a HUGE number to see where new_cases die out
   
  p_s_i <- beta*i_t[t-1]/N
  
   # INCLUDE TIME BEARING BETA
  
  n_t_s_e <- rbinom(1,s_t[t-1],p_s_i)
  n_t_e_i <- rbinom(1,e_t[t-1],sigma)
  n_t_i_star <- rbinom(1,i_t[t-1],gamma+sigma)
  n_t_i_d_and_n_t_i_r <- rmultinom(1,n_t_i_star,c(sigma/(gamma+sigma),gamma/(gamma+sigma)))
  n_t_i_d <- n_t_i_d_and_n_t_i_r[1]
  n_t_i_r <- n_t_i_d_and_n_t_i_r[2]
  
   s_t[t] <- s_t[t-1] - n_t_s_e 
   e_t[t] <- e_t[t-1] - n_t_e_i + n_t_s_e
   i_t[t] <- i_t[t-1] - n_t_i_r + n_t_e_i - n_t_i_d
   d_t[t] <- d_t[t-1] + n_t_i_d 
   r_t[t] <- r_t[t-1] + n_t_i_r
   new_cases[t] <- n_t_e_i # ask if there is a way to prevent double count since someone + in e compartment would be counted as new E and when moves to I would be counted AGAIN as new I
   
   new_deaths[t] <- n_t_i_d
}  
}
plot(new_cases, type ='l', col='red')
lines(new_deaths, col='black')
#abline() - add lines to timesteps where interventions happened

```